---
title: "Cambodia Ab study"
output:
  html_notebook:
    highlight: haddock
    theme: default
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


# Preamble
```{r preamble}
library(here)
here()

library(tidyverse)
library(mgcv)

# grab some colors for plotting
ggplotcols <- scales::hue_pal()(3)
# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

pcols <- c(ggplotcols[2:3],cmagent,ggplotcols[1],corange)
```

# Load Cambodia measurements

These data are publicly available here:
https://osf.io/2kr8b/

See also:
Arnold et al. Integrated Serologic Surveillance of Population Immunity and Disease Transmission, Emerging Infectious Diseases, 2018; 24(7):1188-1194. https://www.ncbi.nlm.nih.gov/pubmed/29912680

```{r load data, warning=FALSE}
d <- readRDS(here("data","cambodia_serology_public.rds"))

# create age strata
d <- d %>%
  filter(age<=75) %>%
  mutate(agecat = cut(age,breaks=c(14,20,25,30,35,40))) 

# reshape to long format
dl <- d %>%
  select(psuid,age,agecat,nie,t24,pfmsp19,pvmsp19,ttmb) %>%
  gather(key=antigen,value=mfi,-psuid,-age,-agecat) %>%
  mutate(antigen = factor(antigen,levels=c("nie","t24","pfmsp19","pvmsp19","ttmb"))) %>%
  group_by(psuid,antigen) %>%
  arrange(psuid,antigen,age)

dl <- dl %>%
  mutate(logmfi = ifelse(mfi>0,log10(mfi),log10(1)))

# add ROC-based cutoffs, published by Priest et al. 2016 PLOS
# http://journals.plos.org/plosntds/article?id=10.1371/journal.pntd.0004699
dl$roccut <- NA
dl$mixcut <- NA
dl$roccut[dl$antigen %in% "nie"] <- log10(792)
dl$roccut[dl$antigen %in% "t24"] <- log10(486)
dl$roccut[dl$antigen %in% "pfmsp19"] <- log10(343)
dl$roccut[dl$antigen %in% "pvmsp19"] <- log10(196)
dl$roccut[dl$antigen %in% "ttmb"] <- log10(100)

# create formatted antigen labels
dl <- dl %>%
  mutate(antigenf=factor(antigen,levels=c("nie","t24","pfmsp19","pvmsp19","ttmb"),
                         labels=c("S. stercoralis NIE","T. solium T24","P. falciparum MSP-1","P. vivax MSP-1","Tetanus toxoid")))



# define seroprevalence based on ROC cutoffs
dl <- dl %>%
  mutate(seropos = ifelse(logmfi>roccut,1,0))
```


# Age stratified distributions

```{r density plot by age}

table(dl$agecat,dl$antigen)

# stratified by age
pdist <- ggplot(data=dl,aes(x=logmfi,fill=antigenf)) +
  facet_grid(agecat~antigen) +
  geom_density(alpha=0.5,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  # geom_vline(aes(xintercept = mixcut),lty=2) +
  scale_fill_manual(values=pcols)+

  labs(x="log10 luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0))

pdist

# unstratified
pdist2 <- ggplot(data=dl,aes(x=logmfi,fill=antigenf)) +
  facet_grid(~antigenf) +
  geom_density(alpha=0.5,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  scale_fill_manual(values=pcols)+
  # geom_vline(aes(xintercept = mixcut),lty=2) +
  labs(x="log10 luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0))

pdist2

# save png file for presentation
ggsave(filename=here("output","cambodia-ab-dists.png"),plot=pdist2,device="png",width=8,height=2.5)


```


# Age dependent means and seroprevalence

## Function for a simultaneous CI around a spline curve
```{r simultaneous CI}
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}

```

## Select smoothing parameters with CV
```{r smoothing parameter selection by CV}
#----------------------------------
# prep data for spline fits
#----------------------------------
dl <- dl %>% 
  ungroup() %>%
  mutate(psuid=factor(psuid),
         dummy=1)
#----------------------------------
# choose the smoothing parameter
# for the splines, k, 
# based on cross-validated MSE
# pick smallest k where CV MSE is 
# close to its minimum
#----------------------------------
library(SuperLearner)
library(mgcv)
source(here("R","SL.mgcv.R")) # mgcv wrapper for SuperLearner

# k=4
set.seed(123)
sld <- filter(dl,antigen=="nie")
cv_nie <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_nie

# k=4
set.seed(123)
sld <- filter(dl,antigen=="t24")
cv_t24 <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_t24

# k=4
set.seed(123)
sld <- filter(dl,antigen=="pfmsp19")
cv_pfmsp1 <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_pfmsp1

# k=7
set.seed(123)
sld <- filter(dl,antigen=="pvmsp19")
cv_pvmsp1 <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_pvmsp1

# k=6
set.seed(123)
sld <- filter(dl,antigen=="ttmb")
cv_ttmb <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_ttmb
```

## Age dependent means
```{r agecurves mfi}

#----------------------------------
# fit GAM with a spline for age
# include a random effect for cluster
# estimate simultaneous CIs around the curve
# for the prediction data, set the dummy to 0 to 
# zero out all of the random effects
# see posts on Stack Exchange for explanation:
# https://stats.stackexchange.com/questions/131106/predicting-with-random-effects-in-mgcv-gam/131116#131116
# https://stats.stackexchange.com/questions/189384/predicting-mean-smooth-in-gam-with-smooth-by-random-factor-interaction
#----------------------------------

# Strongy NIE
fit_nie <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(psuid,bs="re",by=dummy), 
                      data=filter(dl,antigen=="nie"))
newd <- dl %>% filter(antigen=="nie") %>% mutate(dummy=0)
fit_nieci <- gamCI(m=fit_nie,newdata=newd,nreps=10000)

# Taenia t24
fit_t24 <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(psuid,bs="re",by=dummy), 
                      data=filter(dl,antigen=="t24"))
newd <- dl %>% filter(antigen=="t24") %>% mutate(dummy=0)
fit_t24ci <- gamCI(m=fit_t24,newdata=newd,nreps=10000)

# Pf Malaria MSP-1
fit_pfmsp1 <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(psuid,bs="re",by=dummy), 
                      data=filter(dl,antigen=="pfmsp19"))
newd <- dl %>% filter(antigen=="pfmsp19") %>% mutate(dummy=0)
fit_pfmsp1ci <- gamCI(m=fit_pfmsp1,newdata=newd,nreps=10000)

# Pv Malaria MSP-1
fit_pvmsp1 <- mgcv::gam(logmfi~s(age,bs="cr",k=7) + s(psuid,bs="re",by=dummy), 
                      data=filter(dl,antigen=="pvmsp19"))
newd <- dl %>% filter(antigen=="pvmsp19") %>% mutate(dummy=0)
fit_pvmsp1ci <- gamCI(m=fit_pvmsp1,newdata=newd,nreps=10000)

# Tetanus toxoid
fit_ttmb <- mgcv::gam(logmfi~s(age,bs="cr",k=6) + s(psuid,bs="re",by=dummy), 
                      data=filter(dl,antigen=="ttmb"))
newd <- dl %>% filter(antigen=="ttmb") %>% mutate(dummy=0)
fit_ttmbci <- gamCI(m=fit_ttmb,newdata=newd,nreps=10000)



fit_mfi <- bind_rows(fit_nieci,fit_t24ci,fit_pfmsp1ci,fit_pvmsp1ci,fit_ttmbci)

```

```{r mean mfi by age plot}

pagemfi <- ggplot(data=fit_mfi,aes(x=age,y=fit,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_point(aes(x=age,y=logmfi),alpha=0.2,size=0.2)+
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill="black") +
  geom_line(lwd=0.5,alpha=0.5,color="black") +
  geom_smooth(method="loess",se=FALSE,lwd=0.5,color="black") +
  scale_color_manual(values=pcols)+
  scale_y_continuous(breaks=seq(0,4,by=1))+
  labs(x="age, years",y="log10 Luminex response (MFI-bg)") +
  theme_minimal() +
  theme(legend.position="none")

pagemfi


```

## Age dependent seroprevalence
```{r agecurves seroprev}

# Strongy NIE
fitp_nie <- mgcv::gam(seropos~s(age,bs="cr",k=7) + s(psuid,bs="re",by=dummy), family="binomial", 
                      data=filter(dl,antigen=="nie"))
newd <- dl %>% filter(antigen=="nie") %>% mutate(dummy=0)
fitp_nieci <- gamCI(m=fitp_nie,newdata=newd,nreps=10000)

# Taenia t24
fitp_t24 <- mgcv::gam(seropos~s(age,bs="cr",k=9) + s(psuid,bs="re",by=dummy), family="binomial", 
                      data=filter(dl,antigen=="t24"))
newd <- dl %>% filter(antigen=="t24") %>% mutate(dummy=0)
fitp_t24ci <- gamCI(m=fitp_t24,newdata=newd,nreps=10000)

# Pf Malaria MSP-1
fitp_pfmsp1 <- mgcv::gam(seropos~s(age,bs="cr",k=11) + s(psuid,bs="re",by=dummy), family="binomial",
                      data=filter(dl,antigen=="pfmsp19"))
newd <- dl %>% filter(antigen=="pfmsp19") %>% mutate(dummy=0)
fitp_pfmsp1ci <- gamCI(m=fitp_pfmsp1,newdata=newd,nreps=10000)

# Pv Malaria MSP-1
fitp_pvmsp1 <- mgcv::gam(seropos~s(age,bs="cr",k=11) + s(psuid,bs="re",by=dummy), family="binomial",
                      data=filter(dl,antigen=="pvmsp19"))
newd <- dl %>% filter(antigen=="pvmsp19") %>% mutate(dummy=0)
fitp_pvmsp1ci <- gamCI(m=fitp_pvmsp1,newdata=newd,nreps=10000)

# Tetanus
fitp_ttmb <- mgcv::gam(seropos~s(age,bs="cr",k=11) + s(psuid,bs="re",by=dummy), family="binomial",
                      data=filter(dl,antigen=="ttmb"))
newd <- dl %>% filter(antigen=="ttmb") %>% mutate(dummy=0)
fitp_ttmbci <- gamCI(m=fitp_ttmb,newdata=newd,nreps=10000)



fit_seroprev <- bind_rows(fitp_nieci,fitp_t24ci,fitp_pfmsp1ci,fitp_pvmsp1ci,fitp_ttmbci)

# convert linear predictor to prevalance
expitfn <- function(x) {
  exp(x)/(1+exp(x))
}
fit_seroprev <- fit_seroprev %>%
  mutate(fit = expitfn(fit),
         uprP = expitfn(uprP),
         lwrP = expitfn(lwrP),
         uprS = expitfn(uprS),
         lwrS = expitfn(lwrS),
         )

```

```{r seroprev by age plot}

pageprev <- ggplot(data=fit_seroprev,aes(x=age,y=fit,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill="grey50") +
  geom_line(lwd=0.5,alpha=0.5) +
  geom_smooth(method="loess",se=FALSE,lwd=0.5) +
  scale_color_manual(values=pcols)+
  scale_y_continuous(breaks=seq(0,1,by=0.1),labels=seq(0,100,by=10))+
  labs(x="age, years",y="Seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")

pageprev


```
# Cluster level seroprevalence vs. means

```{r cluster means}

dlc <- dl %>%
  group_by(psuid,antigenf) %>%
  mutate(posroc=ifelse(logmfi>roccut,1,0)) %>%
  summarize(n=n(),
            meanmfi=mean(logmfi),
            prevroc=mean(posroc))

dcorr <- dlc %>%
  ungroup() %>%
  group_by(antigenf) %>%
  mutate(corxy=cor(meanmfi,prevroc,method="pearson") ) %>%
  slice(1)

pmfivprev <- ggplot(data=dlc,aes(x=meanmfi,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_point(aes(y=prevroc),alpha=0.7) +
  geom_smooth(aes(y=prevroc),method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dcorr,
            aes(x=2,y=1,label=paste("r ==",sprintf("%1.2f",corxy)) ),
            parse=TRUE,col="grey30") +
  scale_color_manual(values=pcols)+
  scale_y_continuous(breaks=seq(0,1,by=0.1),labels=seq(0,100,by=10))+
  coord_cartesian(ylim=c(0,1),xlim=c(1,4.5)) +
  labs(x="cluster mean log10 Luminex response (MFI-bg)",y="cluster seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")
pmfivprev


# save png file for presentation
ggsave(filename=here("output","cambodia-mfi-v-seroprev.png"),plot=pmfivprev,device="png",width=8,height=2.5)


```
