---
title: "Mbita Kenya Ab study"
output:
  html_notebook:
    highlight: haddock
    theme: default
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


# Preamble
```{r preamble}
library(here)
here()

library(tidyverse)
library(mgcv)
library(gridExtra)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# grab some colors for plotting
ggplotcols <- scales::hue_pal()(3)
# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

pcols <- c(ggplotcols,corange)

```

# Load coastal Kenya measurements
Load AsHb measurements and filter 

```{r load data, warning=FALSE}
d <- readRDS(here("data","mbita_psac.rds"))

# create age strata
d <- d %>%
  select(year,community=vid,
         pid,age=agey,
         wat01,toi01,
         ashb,sea,msp1,vsp5) %>%
  mutate(agecat = cut(age,breaks=c(0,1,2,3,4,6),labels=c("<1 year","1 year","2 years","3 years","4 years")))

# reshape to long format
dl <- d %>%
  select(year,community,wat01,toi01,age,agecat,ashb,sea,msp1,vsp5) %>%
  gather(key=antigen,value=mfi,-year,-community,-wat01,-toi01,-age,-agecat) %>%
  mutate(antigen = factor(antigen,levels=c("ashb","sea","msp1","vsp5"))) %>%
  group_by(year,community,antigen) %>%
  arrange(year,community,antigen,age)

dl <- dl %>%
  mutate(logmfi = ifelse(mfi>0,log10(mfi),log10(1)))

# add ROC-based cutoffs
dl$roccut <- NA
dl$roccut[dl$antigen %in% "ashb"] <- log10(418)
dl$roccut[dl$antigen %in% "sea"] <- log10(965)
dl$roccut[dl$antigen %in% "msp1"] <- log10(170)
dl$roccut[dl$antigen %in% "vsp5"] <- log10(281)

# define seroprevalence based on ROC cutoffs
dl <- dl %>%
  mutate(seropos = ifelse(logmfi>roccut,1,0))


# create formatted antigen labels
dl <- dl %>%
  mutate(antigenf=factor(antigen,levels=c("ashb","sea","msp1","vsp5"),
                         labels=c("Ascaris spp. AsHb",
                                  "S. mansoni SEA",
                                  "P. falciparum MSP-1",
                                  "Giardia spp. VSP-5")))


```


# Age stratified distributions

```{r density plot by age}

table(dl$agecat,dl$antigenf)

pdist <- ggplot(data=dl,aes(x=logmfi,fill=antigen)) +
  facet_grid(agecat~antigenf) +
  geom_density(alpha=0.5,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  scale_fill_manual(values=pcols)+
  # geom_vline(aes(xintercept = mixcut),lty=2) +
  labs(x="log10 luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0))

pdist

# limit to just SEA and Giardia for presentation
pdist2 <- ggplot(data=filter(dl,antigen%in%c("sea","vsp5")),aes(x=logmfi,fill=antigen)) +
  facet_grid(agecat~antigenf) +
  geom_density(alpha=0.75,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  scale_fill_manual(values=pcols[c(2,4)])+
  # geom_vline(aes(xintercept = mixcut),lty=2) +
  labs(x="log10 luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0,size=14),
        strip.text.x=element_text(size=14))

pdist2
# save png file for presentation
ggsave(filename=here("output","mbita-ab-dists-sea-vsp.png"),plot=pdist2,device="png",width=5,height=7)



pdist3 <- ggplot(data=dl,aes(x=logmfi,fill=antigenf)) +
  facet_grid(.~antigenf) +
  geom_density(alpha=0.5,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  scale_fill_manual(values=pcols)+
  # geom_vline(aes(xintercept = mixcut),lty=2) +
  labs(x="log10 luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0))

pdist3

# save png file for presentation
ggsave(filename=here("output","mbita-ab-dists.png"),plot=pdist3,device="png",width=8,height=2)

```


# Age dependent means and seroprevalence

## Function for a simultaneous CI around a spline curve
```{r simultaneous CI}
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}

```

## Select smoothing parameters with CV
```{r smoothing parameter selection by CV}
#----------------------------------
# prep data for spline fits
#----------------------------------
dl <- dl %>% 
  ungroup() %>%
  mutate(community=factor(community),
         dummy=1)
#----------------------------------
# choose the smoothing parameter
# for the splines, k, 
# based on cross-validated MSE
# pick smallest k where CV MSE is 
# close to its minimum
#----------------------------------
# library(SuperLearner)
# library(mgcv)
# source(here("R","SL.mgcv.R")) # mgcv wrapper for SuperLearner
# # k=4
# set.seed(123)
# sld <- filter(dl,antigen=="msp1")
# cv_msp1 <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:10,sep=""))
# cv_msp1
# # k=4
# set.seed(123)
# sld <- filter(dl,antigen=="vsp5")
# cv_vsp5 <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:10,sep=""))
# cv_vsp5
# # k=4
# set.seed(123)
# sld <- filter(dl,antigen=="ashb")
# cv_ashb <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:10,sep=""))
# cv_ashb
# # k=4
# set.seed(123)
# sld <- filter(dl,antigen=="sea")
# cv_sea <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:10,sep=""))
# cv_sea
```

## Age dependent means
```{r agecurves mfi}

#----------------------------------
# fit GAM with a spline for age
# include a random effect for cluster
# estimate simultaneous CIs around the curve
# for the prediction data, set the dummy to 0 to 
# zero out all of the random effects
# see posts on Stack Exchange for explanation:
# https://stats.stackexchange.com/questions/131106/predicting-with-random-effects-in-mgcv-gam/131116#131116
# https://stats.stackexchange.com/questions/189384/predicting-mean-smooth-in-gam-with-smooth-by-random-factor-interaction
#----------------------------------


# Pf Malaria MSP1 
fit_msp1 <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="msp1"))
newd <- dl %>% filter(antigen=="msp1") %>% mutate(dummy=0)
fit_msp1ci <- gamCI(m=fit_msp1,newdata=newd,nreps=10000)

# Giardia VSP5
fit_vsp5 <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="vsp5"))
newd <- dl %>% filter(antigen=="vsp5") %>% mutate(dummy=0)
fit_vsp5ci <- gamCI(m=fit_vsp5,newdata=newd,nreps=10000)

# Ascaris AsHb
fit_ashb <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="ashb"))
newd <- dl %>% filter(antigen=="ashb") %>% mutate(dummy=0)
fit_ashbci <- gamCI(m=fit_ashb,newdata=newd,nreps=10000)

# Schisto SEA
fit_sea <- mgcv::gam(logmfi~s(age,bs="cr",k=4) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="sea"))
newd <- dl %>% filter(antigen=="sea") %>% mutate(dummy=0)
fit_seaci <- gamCI(m=fit_sea,newdata=newd,nreps=10000)

fit_mfi <- bind_rows(fit_msp1ci,fit_vsp5ci,fit_ashbci,fit_seaci)

```

```{r mean mfi by age plot}

pagemfi <- ggplot(data=fit_mfi,aes(x=age,y=fit,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_point(aes(x=age,y=logmfi),alpha=0.1,size=0.2)+
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill="black") +
  geom_line(lwd=0.5,alpha=0.5,color="black") +
  geom_smooth(method="loess",se=FALSE,lwd=0.5,color="black") +
  scale_color_manual(values=pcols)+
  scale_y_continuous(breaks=seq(0,4,by=1))+
  labs(x="age, years",y="log10 Luminex Response (MFI-bg)") +
  theme_minimal() +
  theme(legend.position="none")

pagemfi


```

## Age dependent seroprevalence
```{r agecurves seroprev}



# Ascaris AsHb
fitp_ashb <- mgcv::gam(seropos~s(age,bs="cr",k=11) + s(community,bs="re",by=dummy), family="binomial",
                      data=filter(dl,antigen=="ashb"))
newd <- dl %>% filter(antigen=="ashb") %>% mutate(dummy=0)
fitp_ashbci <- gamCI(m=fitp_ashb,newdata=newd,nreps=10000)

# Schisto SEA
fitp_sea <- mgcv::gam(seropos~s(age,bs="cr",k=9) + s(community,bs="re",by=dummy), family="binomial", 
                      data=filter(dl,antigen=="sea"))
newd <- dl %>% filter(antigen=="sea") %>% mutate(dummy=0)
fitp_seaci <- gamCI(m=fitp_sea,newdata=newd,nreps=10000)

# Pf Malaria MSP1 
fitp_msp1 <- mgcv::gam(seropos~s(age,bs="cr",k=4) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="msp1"))
newd <- dl %>% filter(antigen=="msp1") %>% mutate(dummy=0)
fitp_msp1ci <- gamCI(m=fitp_msp1,newdata=newd,nreps=10000)

# Giardia VSP5
fitp_vsp5 <- mgcv::gam(seropos~s(age,bs="cr",k=4) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="vsp5"))
newd <- dl %>% filter(antigen=="vsp5") %>% mutate(dummy=0)
fitp_vsp5ci <- gamCI(m=fitp_vsp5,newdata=newd,nreps=10000)


fit_seroprev <- bind_rows(fitp_ashbci,fitp_seaci,fitp_msp1ci,fitp_vsp5ci)

# convert linear predictor to prevalance
expitfn <- function(x) {
  exp(x)/(1+exp(x))
}
fit_seroprev <- fit_seroprev %>%
  mutate(fit = expitfn(fit),
         uprP = expitfn(uprP),
         lwrP = expitfn(lwrP),
         uprS = expitfn(uprS),
         lwrS = expitfn(lwrS),
         )

```

```{r seroprev by age plot}

pageprev <- ggplot(data=fit_seroprev,aes(x=age,y=fit,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill="grey50") +
  geom_line(lwd=0.5,alpha=0.5) +
  geom_smooth(method="loess",se=FALSE,lwd=0.5) +
  scale_color_manual(values=pcols)+
  scale_y_continuous(breaks=seq(0,0.8,by=0.1),labels=seq(0,80,by=10))+
  labs(x="age, years",y="Seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")

pageprev


```
# Cluster level seroprevalence vs. means

```{r cluster means}

dlc <- dl %>%
  group_by(community,antigenf) %>%
  mutate(posroc=ifelse(logmfi>roccut,1,0)) %>%
  summarize(n=n(),
            meanmfi=mean(logmfi),
            prevroc=mean(posroc))

# estimate pearson correlation
dcorr <- dlc %>%
  ungroup() %>%
  group_by(antigenf) %>%
  mutate(corroc=cor(meanmfi,prevroc,method="pearson")) %>%
  slice(1)

# estimate regression slope
foreach(ai=levels(dlc$antigenf)) %do% {
  di <- dlc %>% filter(antigenf==ai)
  fit <- lm(prevroc~meanmfi,data=di)
  summary(fit)
}
```


```{r prev vs mfi figure}
pmfivprev <- ggplot(data=dlc,aes(x=meanmfi,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_point(aes(y=prevroc),alpha=0.7) +
  geom_smooth(aes(y=prevroc),method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dcorr,
            aes(x=2.5,y=0.95,label=paste("r ==",sprintf("%1.2f",corroc)) ),
            parse=TRUE,col="grey30") +
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=seq(0,100,by=20))+
  scale_color_manual(values=pcols)+
  coord_cartesian(ylim=c(0,1),xlim=c(1,4.5)) +
  labs(x="community mean log10 MFI",y="community seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")
pmfivprev

# save png file for presentation
ggsave(filename=here("output","mbita-mfi-v-seroprev.png"),plot=pmfivprev,device="png",width=8,height=2.5)


```

# Comparison with FOI

## Constant rate model
Estimate community-level FOI for P. falciparum S. mansoni based on the seroconversion rate (incidence among susceptibles). Assume a constant rate model (i.e., average over all ages).  If we assume a constant force of infection ($\lambda$), equivalent to assuming an exponential model, then it can be shown that a generalized linear model with a complementary log-log link fit with current status, age-prevalence data is equivalent to an exponential proportional hazards model (*Jewell and van der Laan 1995*). 

$\log - \log(1-P(Y|A,W)) = \log \lambda + \log A + \beta W$

Moroever, this model is also equivalent to a catalytic, SIR model with a single, constant rate parameter (*Hens et al. 2010*; *Hens et al. 2012*).

```{r  foi estimation for msp and sea}

# foi_ests <- foreach(ai=c("S. mansoni SEA","P. falciparum MSP-1"),.combine=rbind) %:%
foi_ests <- foreach(ai=levels(dl$antigenf),.combine=rbind) %:%
  foreach(comi=levels(dl$community),.combine=rbind) %do% {
  
  di <- dl %>% filter(antigenf==ai & community==comi)
  gfit <- glm(seropos~1,offset=log(age),data=di,family=binomial(link="cloglog"))
  gsum <- summary(gfit)
  lambda <- as.numeric(exp(gfit$coefficients))
  log_lambda_se  <- sqrt(gsum$cov.unscaled)
  lambda_lb <- as.numeric(exp(gfit$coefficients - 1.96*log_lambda_se))
  lambda_ub <- as.numeric(exp(gfit$coefficients + 1.96*log_lambda_se))
  res <- data.frame(community=comi,antigenf=ai,lambda,lambda_lb,lambda_ub)
  return(res)
}

foi_ests

```



```{r merge foi estimates}
# merge FOI estimates to the cluster-level means
d_foi <- dlc %>%
  # filter(antigenf=="S. mansoni SEA"|antigenf=="P. falciparum MSP-1") %>%
  left_join(foi_ests,by=c("community","antigenf"))# %>%
  # mutate(antigenf=factor(antigenf,levels=c("S. mansoni SEA","P. falciparum MSP-1")))


# estimate spearman correlation
dfoicorr <- d_foi %>%
  ungroup() %>%
  group_by(antigenf) %>%
  mutate(cormfi=cor(meanmfi,lambda,method="spearman"),
         corprev=cor(prevroc,lambda,method="spearman")) %>%
  slice(1)

```

Figure of results
```{r foi vs mfi figure}
pfoivmfi <- ggplot(data=d_foi,aes(x=meanmfi,y=lambda,color=antigenf)) +
  facet_grid(~antigenf)+
  geom_point(,alpha=0.7) +
  # geom_smooth(method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_smooth(method="loess",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dfoicorr,
            aes(x=2.5,y=0.6,label=paste("rho ==",sprintf("%1.2f",cormfi)) ),
            parse=TRUE,col="grey30") +
  scale_x_continuous(breaks=2:4)+
  scale_color_manual(values=pcols)+
  coord_cartesian(xlim=c(1.5,4),ylim=c(0,0.7)) +
  labs(x="community mean log10 MFI",y=expression(paste("community seroconversion rate (",lambda,")"))) +
  theme_minimal() +
  theme(legend.position="none")
pfoivmfi

# save png file for presentation
ggsave(filename=here("output","mbita-mfi-v-foi.png"),plot=pfoivmfi,device="png",width=8,height=2.5)

```
```{r foi vs prev figure}
pfoivprev <- ggplot(data=d_foi,aes(x=prevroc,y=lambda,color=antigenf)) +
  facet_grid(~antigenf)+
  geom_point(alpha=0.7) +
  # geom_smooth(method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_smooth(method="loess",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dfoicorr,
            aes(x=0.2,y=0.6,label=paste("rho ==",sprintf("%1.2f",corprev)) ),
            parse=TRUE,col="grey30") +
  scale_x_continuous(breaks=seq(0,1,by=0.2),labels=seq(0,100,by=20))+
  scale_color_manual(values=pcols)+
  coord_cartesian(xlim=c(0,1),ylim=c(0,0.7)) +
  labs(x="community seroprevalence (%)",y=expression(paste("community seroconversion rate (",lambda,")"))) +
  theme_minimal() +
  theme(legend.position="none")
pfoivprev

# save png file for presentation
ggsave(filename=here("output","mbita-prev-v-foi.png"),plot=pfoivprev,device="png",width=8,height=3)
```

# Effect of smaller samples

Downsample the observations within each community to see how that influences the relationships

Note: this analysis probably won't contribute to the presentation because not sufficient time to think about it and interpret it. 

```{r community downsample}

# estimate means with sample sizes of betweeen 10 and 50 observations
set.seed(123)
ssests <- foreach(ssi=seq(100,10,by=-10),.combine=rbind) %:% 
  foreach(iteri=1:1000,.combine=rbind) %dopar% {
    di <- dl %>%
    group_by(community,antigenf) %>%
    sample_n(size=ssi,replace=TRUE) %>%
    summarize(meanmfi=mean(logmfi),
              seroprev=mean(seropos)) %>%
    mutate(ss=ssi,iter=iteri)
}

# merge in the means with the full sample
dlc2 <- dlc %>%
  select(community,antigenf,n,mean=meanmfi,serop=prevroc) %>%
  left_join(select(d_foi,community,antigenf,lambda),by=c("community","antigenf")) %>%
  mutate(antigenf=factor(antigenf,levels=levels(dl$antigenf)))
ssests2 <- left_join(ssests,dlc2,by=c("community","antigenf"))


# average over bootstrap estimates
ssmeans <- ssests2 %>%
  group_by(community,antigenf,ss) %>%
  summarize(mean=mean(mean),
            mu=mean(meanmfi),
            mu_sd=sd(meanmfi),
            mu_lb=quantile(meanmfi,probs=c(0.025)),
            mu_ub=quantile(meanmfi,probs=c(0.975)),
            mu_mse=mean((meanmfi-mean)^2),
            mu_bias=mean(meanmfi-mean),
            
            serop=mean(serop),
            prev=mean(seroprev),
            prev_sd=sd(seroprev),
            prev_lb=quantile(seroprev,probs=c(0.025)),
            prev_ub=quantile(seroprev,probs=c(0.975)),
            prev_mse=mean((seroprev-serop)^2),
            prev_bias=mean(seroprev-serop)
            )

# correlation betwen community-level means and seroprev, over bootstrap estimates
sscorr <- ssests2 %>%
  group_by(antigenf,ss,iter) %>%
  summarize(cor_mean=cor(meanmfi,serop,method="pearson"),
            cor_prev=cor(seroprev,serop,method="pearson")) %>%
  ungroup() %>%
  group_by(antigenf,ss) %>%
  summarize(cor_mu=mean(cor_mean),
            cor_mu_lb=quantile(cor_mean,probs=c(0.025)),
            cor_mu_ub=quantile(cor_mean,probs=c(0.975)),
            cor_pr=mean(cor_prev),
            cor_pr_lb=quantile(cor_prev,probs=c(0.025)),
            cor_pr_ub=quantile(cor_prev,probs=c(0.975)),
            )

# correlation betwen community-level means and seroprev, over bootstrap estimates
sscorr_foi <- ssests2 %>%
  group_by(antigenf,ss,iter) %>%
  filter(!is.na(lambda)) %>%
  summarize(cor_mean=cor(meanmfi,lambda,method="spearman"),
            cor_prev=cor(seroprev,lambda,method="spearman")) %>%
  ungroup() %>%
  group_by(antigenf,ss) %>%
  summarize(cor_mu=mean(cor_mean),
            cor_mu_lb=quantile(cor_mean,probs=c(0.025)),
            cor_mu_ub=quantile(cor_mean,probs=c(0.975)),
            cor_pr=mean(cor_prev),
            cor_pr_lb=quantile(cor_prev,probs=c(0.025)),
            cor_pr_ub=quantile(cor_prev,probs=c(0.975)),
            )

```
Plot a figure of RMSE in community-level estimates by sample size.
```{r downsample rmse figs}
ssmumsep <- ggplot(data=ssmeans,aes(x=ss,y=sqrt(mu_mse),group=community)) +
  facet_grid(.~antigenf)+
  geom_line(alpha=0.2) +
  theme_minimal()
ssmumsep

ssprevmsep <- ggplot(data=ssmeans,aes(x=ss,y=sqrt(prev_mse),group=community)) +
  facet_grid(.~antigenf)+
  geom_line(alpha=0.2) +
  theme_minimal()
ssprevmsep

```

Plot a figure of the correlation between the seroprevalence estimated in the full sample versus cluster level mean (blue) or cluster level seroprevalence (orange) estimated at each smaller sample size

```{r downsample corr figure}
sscorp <- ggplot(data=sscorr,aes(x=ss)) +
  facet_grid(.~antigenf)+
  geom_ribbon(aes(ymin=cor_mu_lb,ymax=cor_mu_ub),color=NA,fill=cblue,alpha=0.2)+
  geom_line(aes(y=cor_mu),color=cblue) +
  geom_ribbon(aes(ymin=cor_pr_lb,ymax=cor_pr_ub),color=NA,fill=corange,alpha=0.2)+
  geom_line(aes(y=cor_pr),color=corange)+
  coord_cartesian(ylim=c(0,1))+
  labs(y="Correlation with community level seroprevalence",x="Sample size per community") +
  theme_minimal()
sscorp

```

Plot a figure of the correlation between the seroconversion rate estimated in the full sample versus cluster level mean (blue) or cluster level seroprevalence (orange) estimated at each smaller sample size

```{r downsample corr foi figure}
sscorfoip <- ggplot(data=sscorr_foi,aes(x=ss)) +
  facet_grid(.~antigenf)+
  geom_ribbon(aes(ymin=cor_mu_lb,ymax=cor_mu_ub),color=NA,fill=cblue,alpha=0.2)+
  geom_line(aes(y=cor_mu),color=cblue) +
  geom_ribbon(aes(ymin=cor_pr_lb,ymax=cor_pr_ub),color=NA,fill=corange,alpha=0.2)+
  geom_line(aes(y=cor_pr),color=corange)+
  coord_cartesian(ylim=c(0,1))+
  labs(y="Correlation with community level seroconversion rate",x="Sample size per community") +
  theme_minimal()
sscorfoip

```

Based on these figures, there is no loss of information from reducing quantitative measures to seroprevalence, and if anything seroprevalence performs slightly better with respect to correlation measures.





