---
title: "Coastal Kenya Ab study"
output:
  html_notebook:
    highlight: haddock
    theme: default
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


# Preamble
```{r preamble}
library(here)
here()

library(tidyverse)
library(foreach)
library(mgcv)
```

# Load coastal Kenya measurements
Load AsHb measurements and filter 

```{r load data, warning=FALSE}
d <- readRDS(here("data","coastal_kenya.rds"))

# create age strata
d <- d %>%
  filter(age<=75) %>%
  mutate(agecat = cut(age,breaks=c(0,5,10,20,30,40,75))) %>%
  rename(ashb=ascaris)

# reshape to long format
dl <- d %>%
  select(community,age,agecat,sea,nie,ashb) %>%
  gather(key=antigen,value=mfi,-community,-age,-agecat) %>%
  group_by(community,antigen) %>%
  arrange(community,antigen,age)

dl <- dl %>%
  mutate(logmfi = ifelse(mfi>0,log10(mfi),log10(1)))

# create formatted antigen labels
dl <- dl %>%
  mutate(antigenf=factor(antigen,levels=c("ashb","sea","nie"),
                         labels=c("Ascaris spp. AsHb","S. mansoni SEA","S. stercoralis NIE")))


# add ROC and mixture model-based cutoffs
dl$roccut <- NA
dl$mixcut <- NA
dl$roccut[dl$antigen %in% "sea"] <- log10(965)
dl$mixcut[dl$antigen %in% "sea"] <- log10(476)
dl$roccut[dl$antigen %in% "nie"] <- log10(628)
dl$mixcut[dl$antigen %in% "nie"] <- log10(645)
dl$roccut[dl$antigen %in% "ashb"] <- log10(386)
dl$mixcut[dl$antigen %in% "ashb"] <- log10(780)

# define seroprevalence based on ROC cutoffs
dl <- dl %>%
  mutate(seropos = ifelse(logmfi>roccut,1,0))
```


# Age stratified distributions

```{r density plot ashb}
ggplotcols <- scales::hue_pal()(3)
pdistall <- ggplot(data=dl,aes(x=logmfi,fill=antigenf)) +
  facet_grid(.~antigenf) +
  geom_density(alpha=0.5,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  # scale_fill_manual(values=ggplotcols[1])+
  scale_x_continuous(breaks=0:4,labels=c(expression(10^0),expression(10^1),expression(10^2),expression(10^3),expression(10^4))) +
  labs(x="luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0))

pdistall

# save png file for presentation
ggsave(filename=here("output","coastal-ab-dists.png"),plot=pdistall,device="png",width=6,height=2.5)

```

```{r density plot by age}

table(dl$agecat,dl$antigen)

pdist <- ggplot(data=dl,aes(x=logmfi,fill=antigenf)) +
  facet_grid(agecat~antigenf) +
  geom_density(alpha=0.5,color=NA) +
  geom_vline(aes(xintercept = roccut),lty=1) +
  # geom_vline(aes(xintercept = mixcut),lty=2) +
  labs(x="log10 luminex response (MFI-bg)") +
  theme_minimal() + 
  theme(legend.position = "none",
        strip.text.y=element_text(angle=0))

pdist

```


# Age dependent means and seroprevalence

## Function for a simultaneous CI around a spline curve
```{r simultaneous CI}
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}

```

## Select smoothing parameters with CV
```{r smoothing parameter selection by CV}
#----------------------------------
# prep data for spline fits
#----------------------------------
dl <- dl %>% 
  ungroup() %>%
  mutate(community=factor(community),
         dummy=1)
#----------------------------------
# choose the smoothing parameter
# for the splines, k, 
# based on cross-validated MSE
# pick smallest k where CV MSE is 
# close to its minimum
#----------------------------------
library(SuperLearner)
library(mgcv)
source(here("R","SL.mgcv.R")) # mgcv wrapper for SuperLearner
set.seed(123)
sld <- filter(dl,antigen=="ashb")
cv_ashb <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_ashb
# k=11
set.seed(123)
sld <- filter(dl,antigen=="nie")
cv_nie <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_nie
# k=7
set.seed(123)
sld <- filter(dl,antigen=="sea")
cv_sea <- SuperLearner(Y=sld$logmfi,X=select(sld,age), SL.library = paste("SL.mgcv.k",4:20,sep=""))
cv_sea
# k=9
```

## Age dependent means
```{r agecurves mfi}

#----------------------------------
# fit GAM with a spline for age
# include a random effect for cluster
# estimate simultaneous CIs around the curve
# for the prediction data, set the dummy to 0 to 
# zero out all of the random effects
# see posts on Stack Exchange for explanation:
# https://stats.stackexchange.com/questions/131106/predicting-with-random-effects-in-mgcv-gam/131116#131116
# https://stats.stackexchange.com/questions/189384/predicting-mean-smooth-in-gam-with-smooth-by-random-factor-interaction
#----------------------------------

# Ascaris AsHb
fit_ashb <- mgcv::gam(logmfi~s(age,bs="cr",k=11) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="ashb"))
newd <- dl %>% filter(antigen=="ashb") %>% mutate(dummy=0)
fit_ashbci <- gamCI(m=fit_ashb,newdata=newd,nreps=10000)


# Schisto SEA
fit_sea <- mgcv::gam(logmfi~s(age,bs="cr",k=9) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="sea"))
newd <- dl %>% filter(antigen=="sea") %>% mutate(dummy=0)
fit_seaci <- gamCI(m=fit_sea,newdata=newd,nreps=10000)

# Strongy NIE
fit_nie <- mgcv::gam(logmfi~s(age,bs="cr",k=7) + s(community,bs="re",by=dummy), 
                      data=filter(dl,antigen=="nie"))
newd <- dl %>% filter(antigen=="nie") %>% mutate(dummy=0)
fit_nieci <- gamCI(m=fit_nie,newdata=newd,nreps=10000)

fit_mfi <- bind_rows(fit_ashbci,fit_seaci,fit_nieci)

```

```{r mean mfi by age plot}

pagemfi <- ggplot(data=fit_mfi,aes(x=age,y=fit,color=antigen)) +
  facet_grid(~antigen) +
  geom_point(aes(x=age,y=logmfi),alpha=0.1,size=0.2)+
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill="black") +
  geom_line(lwd=0.5,alpha=0.5,color="black") +
  geom_smooth(method="loess",se=FALSE,lwd=0.5,color="black") +
  scale_y_continuous(breaks=seq(0,4,by=1))+
  labs(x="age, years",y="log10 Luminex Response (MFI-bg)") +
  theme_minimal() +
  theme(legend.position="none")

pagemfi


```

## Age dependent seroprevalence
```{r agecurves seroprev}

# Ascaris AsHb
fitp_ashb <- mgcv::gam(seropos~s(age,bs="cr",k=11) + s(community,bs="re",by=dummy), family="binomial",
                      data=filter(dl,antigen=="ashb"))
newd <- dl %>% filter(antigen=="ashb") %>% mutate(dummy=0)
fitp_ashbci <- gamCI(m=fitp_ashb,newdata=newd,nreps=10000)

# Schisto SEA
fitp_sea <- mgcv::gam(seropos~s(age,bs="cr",k=9) + s(community,bs="re",by=dummy), family="binomial", 
                      data=filter(dl,antigen=="sea"))
newd <- dl %>% filter(antigen=="sea") %>% mutate(dummy=0)
fitp_seaci <- gamCI(m=fitp_sea,newdata=newd,nreps=10000)

# Strongy NIE
fitp_nie <- mgcv::gam(seropos~s(age,bs="cr",k=7) + s(community,bs="re",by=dummy), family="binomial", 
                      data=filter(dl,antigen=="nie"))
newd <- dl %>% filter(antigen=="nie") %>% mutate(dummy=0)
fitp_nieci <- gamCI(m=fitp_nie,newdata=newd,nreps=10000)

fit_seroprev <- bind_rows(fitp_ashbci,fitp_seaci,fitp_nieci)

# convert linear predictor to prevalance
expitfn <- function(x) {
  exp(x)/(1+exp(x))
}
fit_seroprev <- fit_seroprev %>%
  mutate(fit = expitfn(fit),
         uprP = expitfn(uprP),
         lwrP = expitfn(lwrP),
         uprS = expitfn(uprS),
         lwrS = expitfn(lwrS),
         )

```

```{r seroprev by age plot}

pageprev <- ggplot(data=fit_seroprev,aes(x=age,y=fit,color=antigen)) +
  facet_grid(~antigen) +
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill="grey50") +
  geom_line(lwd=0.5,alpha=0.5) +
  geom_smooth(method="loess",se=FALSE,lwd=0.5) +
  scale_y_continuous(breaks=seq(0,0.8,by=0.1),labels=seq(0,80,by=10))+
  labs(x="age, years",y="Seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")

pageprev


```
# Cluster level seroprevalence vs. means

```{r cluster means}

dlc <- dl %>%
  group_by(community,antigenf) %>%
  mutate(posmix=ifelse(logmfi>mixcut,1,0),
         posroc=ifelse(logmfi>roccut,1,0)) %>%
  summarize(n=n(),
            meanmfi=mean(logmfi),
            prevroc=mean(posroc),
            prevmix=mean(posmix))

dcorr <- dlc %>%
  ungroup() %>%
  group_by(antigenf) %>%
  mutate(corroc=cor(meanmfi,prevroc,method="pearson"),
         cormix=cor(meanmfi,prevmix,method="pearson")) %>%
  slice(1)

pmfivprev <- ggplot(data=dlc,aes(x=meanmfi,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_point(aes(y=prevroc),alpha=0.7) +
  geom_smooth(aes(y=prevroc),method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dcorr,
            aes(x=3,y=0.1,label=paste("r ==",sprintf("%1.2f",corroc)) ),
            parse=TRUE,col="grey30",hjust=0) +
  scale_y_continuous(breaks=seq(0,0.8,by=0.1),labels=seq(0,80,by=10))+
  coord_cartesian(ylim=c(0,0.8),xlim=c(2,3.75)) +
  labs(x="community mean log10 MFI",y="community seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")
pmfivprev

# save png file for presentation
ggsave(filename=here("output","coastal-mfi-v-seroprev.png"),plot=pmfivprev,device="png",width=6,height=2.5)

# include estimates using mixture model based cutoff
pmfivprev2 <- ggplot(data=dlc,aes(x=meanmfi,color=antigenf)) +
  facet_grid(~antigenf) +
  geom_point(aes(y=prevroc),alpha=0.7) +
  geom_smooth(aes(y=prevroc),method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dcorr,
            aes(x=2.5,y=0.1,label=paste("r_roc ==",sprintf("%1.2f",corroc)) ),
            parse=TRUE,col="grey30",hjust=0) +
  
  geom_point(aes(y=prevmix),shape=21) +
  geom_smooth(aes(y=prevmix),method="glm",color="gray40",lty=2,lwd=0.5,se=FALSE) +
  geom_text(data=dcorr,
            aes(x=2.5,y=0.05,label=paste("r_mix ==",sprintf("%1.2f",cormix)) ),
            parse=TRUE,col="grey30",hjust=0) +
  
  scale_y_continuous(breaks=seq(0,0.8,by=0.1),labels=seq(0,80,by=10))+
  coord_cartesian(ylim=c(0,0.8),xlim=c(2,3.75)) +
  labs(x="community mean log10 MFI",y="community seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position="none")
pmfivprev2
```
# Comparison with FOI

## Constant rate model
Estimate community-level FOI for P. falciparum S. mansoni based on the seroconversion rate (incidence among susceptibles). Assume a constant rate model (i.e., average over all ages).  If we assume a constant force of infection ($\lambda$), equivalent to assuming an exponential model, then it can be shown that a generalized linear model with a complementary log-log link fit with current status, age-prevalence data is equivalent to an exponential proportional hazards model (*Jewell and van der Laan 1995*). 

$\log - \log(1-P(Y|A,W)) = \log \lambda + \log A + \beta W$

Moroever, this model is also equivalent to a catalytic, SIR model with a single, constant rate parameter (*Hens et al. 2010*; *Hens et al. 2012*).

```{r  foi estimation for msp and sea}

foi_ests <- foreach(ai=levels(dl$antigenf),.combine=rbind) %:%
  foreach(comi=levels(dl$community),.combine=rbind) %do% {
  
  di <- dl %>% filter(antigenf==ai & community==comi)
  gfit <- glm(seropos~1,offset=log(age),data=di,family=binomial(link="cloglog"))
  gsum <- summary(gfit)
  lambda <- as.numeric(exp(gfit$coefficients))
  log_lambda_se  <- sqrt(gsum$cov.unscaled)
  lambda_lb <- as.numeric(exp(gfit$coefficients - 1.96*log_lambda_se))
  lambda_ub <- as.numeric(exp(gfit$coefficients + 1.96*log_lambda_se))
  res <- data.frame(community=comi,antigenf=ai,lambda,lambda_lb,lambda_ub)
  return(res)
}
foi_ests <- foi_ests %>% mutate(antigenf=factor(antigenf,levels=levels(dl$antigenf)))
foi_ests

```

```{r merge foi estimates}
# merge FOI estimates to the cluster-level means
d_foi <- dlc %>%
  left_join(foi_ests,by=c("community","antigenf")) %>%
  mutate(antigenf=factor(antigenf,levels=levels(dl$antigenf)))


# estimate spearman correlation
dfoicorr <- d_foi %>%
  ungroup() %>%
  group_by(antigenf) %>%
  mutate(cormfi=cor(meanmfi,lambda,method="pearson"),
         corprev=cor(prevroc,lambda,method="pearson")) %>%
  select(antigenf,cormfi,corprev) %>%
  slice(1)

```


Figure of results
```{r foi vs mfi figure}
pfoivmfi <- ggplot(data=d_foi,aes(x=meanmfi,y=lambda,color=antigenf)) +
  facet_grid(~antigenf)+
  geom_point(,alpha=0.7) +
  geom_smooth(method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dfoicorr,
            aes(x=3,y=0.05,label=paste("r ==",sprintf("%1.2f",cormfi)) ),
            parse=TRUE,col="grey30") +
  scale_x_continuous(breaks=2:4)+
  coord_cartesian(xlim=c(2,4),ylim=c(0,0.05)) +
  labs(x="community mean log10 MFI",y=expression(paste("community seroconversion rate (",lambda,")"))) +
  theme_minimal() +
  theme(legend.position="none")
pfoivmfi
```
```{r foi vs prev figure}
pfoivprev <- ggplot(data=d_foi,aes(x=prevroc,y=lambda,color=antigenf)) +
  facet_grid(~antigenf)+
  geom_point(alpha=0.7) +
  geom_smooth(method="glm",color="gray40",lwd=0.5,se=FALSE) +
  geom_text(data=dfoicorr,
            aes(x=0.5,y=0.05,label=paste("r ==",sprintf("%1.2f",corprev)) ),
            parse=TRUE,col="grey30") +
  scale_x_continuous(breaks=seq(0,1,by=0.2),labels=seq(0,100,by=20))+
  # scale_color_manual(values=pcols[c(3,4)])+
  coord_cartesian(xlim=c(0,1),ylim=c(0,0.05)) +
  labs(x="community seroprevalence (%)",y=expression(paste("community seroconversion rate (",lambda,")"))) +
  theme_minimal() +
  theme(legend.position="none")
pfoivprev


```

